---
title: "data_analysis_4muestras"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load data

```{r}
library(ggplot2)
library(gridExtra)
library(tidyr)
library(plotly)
library(ggrepel)


##CHANGE THIS: the path to load peak matrix - remember! it ends in .zip if generated with rMSIproc, and ends in .pkmat if generated with rMSI2
pks <- rMSI2::LoadPeakMatrix("/home/guest/msidata2/TFG_Claudia/Procesado/analysis4muestras/merged-peakmatrix.pkmat")

tic_data <- pks$intensity/pks$normalizations$TIC

##CHANGE THIS: modifiy "new names" to write the names that you want to see appearing as the name of tha sample. IMPORTANT! The order has to be the same as in pks$names. You can print pks$names in the console to see what it is the original order
new_names <- c("WTHFD_2", "WTHFD_3","KOHFD_3","WTCT_1")


region <- unlist(sapply(1:length(new_names), function(x){rep(new_names[x], pks$numPixels[x])}))
pks$region <- as.factor(region)

categoria <-c("WT", "WT", "KO", "WT")
tipo <- unlist(sapply(1:length(categoria), function(x){rep(categoria[x], pks$numPixels[x])}))
pks$tipo <- as.factor(tipo)

############

#write.csv(pks$names, "/home/guest/msidata2/TFG_Claudia/Procesado/analysis4muestras/nombres.csv", row.names=FALSE)

#new_datos <-read.csv("/home/guest/msidata2/TFG_Claudia/Procesado/analysis4muestras/nombres4.csv", header = FALSE, sep=",")

#new_nombres <- as.vector(new_datos[1]$V1)

#region <- unlist(sapply(1:length(new_nombres), function(x){rep(new_nombres[x], pks$numPixels[x])}))
#pks$region <- as.factor(region)

#categoria <- as.vector(new_datos[2]$V2)


#tipo <- unlist(sapply(1:length(categoria), function(x){rep(categoria[x], pks$numPixels[x])}))
#pks$tipo <- as.factor(tipo)

#########

##CHANGE THIS: the order in which you want the samples to appear. You have to do that by using the names that appear in pks$names
#order <- c("0315_pos_DAN_liverhomo_6A_matrix-only-proc", "0315_neg_DAN_liverhomo_6A_matrix-only-proc",
           #"0315_pos_DAN_liverhomo_6A_10min-proc", "0315_neg_DAN_liverhomo_6A_10min-proc")

##CHANGE THIS: Put the new names in the order in which the samples should appear
#order_names <- c("6,000 A POS (matrix)", "6,000 A NEG (matrix)", "6,000 A POS", "6,000 A NEG")


#rMSI2::plotPeakImage(pks, 500)
```

# TIC accross all samples

```{r echo=FALSE, message=FALSE, warning=FALSE}

##No need to change anything here

df_tic <- data.frame(
  names = pks$region,
  tic = pks$normalizations$TIC
)

bx <- ggplot() + geom_boxplot(aes(x = df_tic$names, y = df_tic$tic)) + xlab("Thickness in nm") + ylab("Total intensity") +
  geom_jitter(aes(x = df_tic$names, y = df_tic$tic), color = "red", alpha = 0.2, size = 0.4) + ggtitle("Total TIC")
bx

rMSIproc::plotValuesImage(pks, pks$normalizations$TIC, labels = new_names, scale_title = "Total TIC in each pixel")
```

# Average spectra of the samples

```{r echo=FALSE, message=FALSE, warning=FALSE}

## CHANGE THIS: This creates an empty data frame. The data frame should have as many columns with "int" as samples you want to compare, plus one for the total intensity (in this examples, there were 6 samples to compare)
df <- data.frame(mass = pks$mass,
                 int_WTHFD_2 = rep(NA, length(pks$mass)),#create as many of these columns as samples you have
                 int_WTHFD_3 = rep(NA, length(pks$mass)),
                 int_KOHFD_3 = rep(NA, length(pks$mass)),
                 int_WTCT_1 = rep(NA, length(pks$mass)),
                 #int_WT= rep(NA, length(pks$mass)),
                 #int_KO= rep(NA, length(pks$mass)),
                 int = rep(NA, length(pks$mass)))

####PREPARAR DATAFRAME PARA VOLCANO
df_tipo<-data.frame(mass=pks$mass,
                 int_WT= rep(NA, length(pks$mass)),
                 int_KO= rep(NA, length(pks$mass)),
                 int = rep(NA, length(pks$mass)))

for( i in 1:length(pks$mass))
{
  df_tipo$int_WT[i] <-mean(pks$intensity[which(pks$tipo=="WT"), i])
  df_tipo$int_KO[i] <-mean(pks$intensity[which(pks$tipo=="KO"), i])
  df_tipo$int[i] <- mean(pks$intensity[i]) 
  
}
#esta columna esta bien igual no? porque hace la media de los 4 que entran en las dos categorias ¿?
######

df_ordenado <- df[,order(names(df))]

######

#CHANGE THIS: This expresion "levels(pks$region)[x]" indicates in which order the sample was in the original peak matrix. In this example, the sample that we want to appear first (40 sec _ 1) appeared in the 6th position in pks$names. Change the "x" in "levels(pks$region)[x]" accordingly
for( i in 1:length(pks$mass))
{
  df$int_KOHFD_3[i] <- mean(pks$intensity[which(pks$region==levels(pks$region)[1]), i]) 
  df$int_WTCT_1[i] <- mean(pks$intensity[which(pks$region==levels(pks$region)[2]), i])
  df$int_WTHFD_2[i] <- mean(pks$intensity[which(pks$region==levels(pks$region)[3]), i]) 
  df$int_WTHFD_3[i] <- mean(pks$intensity[which(pks$region==levels(pks$region)[4]), i])
  #df$int_WT[i] <-mean(pks$intensity[which(pks$tipo=="WT")],i)
  #df$int_KO[i] <-mean(pks$intensity[which(pks$tipo=="KO")],i)
  df$int[i] <- mean(pks$intensity[i])
}

####PRUEBAS AUTOMATIZAR 2
df_auto <-aggregate(pks$region~pks$intensity, data=pks$intensity, mean)

intensity_matrix <- data.frame(pks$intensity)
colnames(intensity_matrix) <- pks$mass
intensity_matrix$region <- pks$region
list_averages <-aggregate(intensity_matrix[, 1:(ncol(intensity_matrix)-1)], list(intensity_matrix$region), mean)
write.csv(list_averages,"the path and the name of the file you want to create")

####PRUEBAS AUTOMATIZAR
cnt<-1
for (i in 1:length(pks$mass))
{
  for(j in 2:(ncol(df_ordenado)-1))
  {
    j
    df_ordenado[i][j]<-mean(pks$intensity[which(pks$region==levels(pks$region)[cnt]), i])
    cnt<-cnt+1
  }
  #df_ordenado$int[i] <- mean(pks$intensity[i])
}

cont<-1
for (j in 2:(ncol(df_ordenado)-1))
  {
  #cat("j es", j, "\n")
  #cat ("contador es ", cont, "\n")
  for (i in 1:length(pks$mass))
  {
    #cat("cont dentro bucle es",cont , "\n")
    df_ordenado[i][j]<-mean(pks$intensity[which(pks$region==levels(pks$region)[cont]), i])
  }
  cont<-cont+1
}

######

df_long<- gather(df, region, intensity, int_WTHFD_2:int_WTCT_1, factor_key=TRUE)
box_names <- c(int_WTHFD_2 = new_names[1], int_WTHFD_3 = new_names[2], int_KOHFD_3= new_names[3], int_WTCT_1=new_names[4])
q <- ggplot(data = df_long) +
  geom_segment(aes(x=df_long$mass, y=df_long$intensity, xend = df_long$mass, yend = 0)) + theme_light() + ylab("Intensity") + xlab("m/z") + facet_grid(df_long$region ~ ., labeller = as_labeller(box_names))
q 
v <- ggplot(data = df_long) +
  geom_segment(aes(x=df_long$mass, y=df_long$intensity, xend = df_long$mass, yend = 0)) + theme_light() + ylab("Intensity") + xlab("m/z") + facet_grid(~df_long$region , labeller = as_labeller(box_names))
v 
#ggplotly(v)


```

#Ion images and boxplots of the intensity of the ten most intense ions (not isotopes)
```{r echo=FALSE, message=FALSE, warning=FALSE}
##No need to change anything here

isotop <- rMSIproc:::isotopeAnnotation(pks, 2, 5, scoreThreshold = 0.8)
monoiso_index <- isotop$monoisotopicPeaks #index de los monoisotopos
monoiso_mass <- pks$mass[monoiso_index] # cogemos las masas de esos index
df_selec_ions <-  df[monoiso_index, ]  #cogemos toda las columnas decada una de las masas 
df_selec_ions <- df_selec_ions[order(df_selec_ions$int, decreasing = TRUE),] #ordenamos de mas intenso a menos

ten_mostintense <- head(df_selec_ions , 10)
ten_mostintense
#ten_lessintense <-tail(df_selec_ions,10) #los ultimos 10
#ten_lessintense
for (i in 1:length(ten_mostintense$mass)){
rMSIproc::plotPeakImage(pks, ten_mostintense$mass[i], labels = new_names)
}

pks$region <- factor(pks$region, levels = new_names)

for (i in 1:length(ten_mostintense$mass)){
f <- ggplot() + geom_boxplot(aes(x=pks$region, y=pks$intensity[, which(pks$mass==ten_mostintense$mass[i])])) +
  geom_jitter(aes(x=pks$region, y=pks$intensity[, which(pks$mass==ten_mostintense$mass[i])]), color = "blue", size = 0.4, alpha = 0.2) +
  ylab("Intensity") +
  xlab("Sample name") +
  ggtitle(paste("Intensity of ion", round(ten_mostintense$mass[i], 4))) +
  theme_light()
print(f)
}
```

##Volcano plot 

```{r echo = FALSE, message = FALSE, warning = FALSE}
#levels(pks$region)
# "8,000 A FRESH"  "8,000 A FROZEN"  


threshold_fc = 2
threshold_p = 0.05

var1_title <- "WT"
var2_title <- "KO"

#### PRUEBAS
#var1  <- pks$intensity[var1_title %in% pks$region , ] #POS matrix

#varKO  <- pks$intensity[pks$region =="KOHFD_3", ]#pks$region == "WTHFD_2"|"WTHFD_3", ] #POS str_view(x,"an")

#varWCT  <- pks$intensity[pks$region =="WTCT_1", ]

#var1  <- pks$intensity[pks$region == "WTHFD_2" , ] 
#varp <- pks$intensity[pks$region == "WTHFD_3" , ]
#varWT<- rbind(var1,varp)  #juntamos las dos muestras de WTHFD en una 

##### SEGUN CATEGORIA
v1 <-pks$intensity[pks$tipo =="WT", ]
v2 <-pks$intensity[pks$tipo =="KO", ]

foldchange_log2 <- function(var1, var2)
{
  mv1 <- mean(var1)
  mv2 <- mean(var2)
  return (log2(mv2/mv1))
}

dfFC <- data.frame (mzID = 1:length(pks$mass),
                    mz = pks$mass,
                    meanint_1 = rep(NA, length(pks$mass)),
                    meanint_2 = rep(NA, length(pks$mass)),
                    FC = rep(NA, length(pks$mass)),
                    pval = rep(NA, length(pks$mass)))

#VOLCANO varWT vs varWCT
for( i in 1:length(pks$mass))
{
  dfFC$meanint_1[i] <- v1[ ,i]
  dfFC$meanint_2[i] <- v2[ ,i]
  dfFC$FC[i] <- foldchange_log2(v1[, i], v2[ , i] )
  dfFC$pval[i] <- kruskal.test(list( v1[, i], v2[ , i]))$p.value
  dfFC$threshold_fc[i] <- (abs(dfFC$FC[i]) >= threshold_fc)
  dfFC$threshold_pval[i] <- dfFC$pval[i] <= threshold_p
  dfFC$threshold[i] <- all(dfFC$threshold_pval[i], dfFC$threshold_fc[i])
 }

sel_dfFC <- dfFC[dfFC$threshold==TRUE, ]


vol <- ggplot(data = dfFC, aes(x=FC, y =-log10(dfFC$pval))) + geom_point(aes(colour=dfFC$threshold, mapping = pks$mass), alpha=0.75, size=1.75) + theme_light() + xlab("fold change") + ylab("-log10 p-value") + ggtitle(paste("Volcano plot:", var1_title,  "(left) vs", var2_title, "(right)")) + scale_color_discrete(name = "Threshold")

vol + geom_text_repel(data = sel_dfFC, aes(x= sel_dfFC$FC, y=-log10(sel_dfFC$pval), label = round(sel_dfFC$mz, 4)), colour = "grey")

s <- ggplot() +
    geom_segment(aes(x=df_tipo$mass, y=df_tipo$int_WT, xend = df_tipo$mass, yend = 0), colour = "red") + 
    geom_segment(aes(x=df_tipo$mass, y=df_tipo$int_KO*(-1), xend = df_tipo$mass, yend = 0), colour = "blue") +
    geom_segment(aes(x=sel_dfFC$mz, y=sel_dfFC$meanint_1, xend = sel_dfFC$mz, yend = 0), colour = "green") + 
    theme_light() + ylab("Intensity") + xlab("m/z") +
    annotate(geom = "text", label = var1_title, color = "red", x = 500, y = max(df_tipo$int_WT)) + 
    annotate(geom = "text", label = var2_title, color = "blue", x = 500, y = min(df_tipo$int_KO * (-1)))
  
s

print("m/z that are significantly different between WTHFD and KOHFD")
knitr::kable(sel_dfFC[,2])

#for (i in 1:length(sel_dfFC$mz)){
#  select <- which(pks$region == var1_title | pks$region == var2_title) 
#f <- ggplot() + geom_boxplot(aes(x=pks$region[select], y=pks$intensity[select, which(pks$mass==sel_dfFC$mz[i])])) +
#  geom_jitter(aes(x=pks$region[select], y=pks$intensity[select, which(pks$mass==sel_dfFC$mz[i])]), color = "blue", size = 0.4, alpha = 0.2) +
#  ylab("Intensity") +
#  xlab("Sample name") +
#  ggtitle(paste("Intensity of ion", round(pks$mass==sel_dfFC$mz[i], 4))) +
#  theme_light()
#print(f)
#}

#for (i in 1:length(sel_dfFC$mz)){
# rMSIproc::plotPeakImage(pks, sel_dfFC$mz[i], labels = new_names)
#}



```

##PCs and Clusters (TIC)

```{r echo=FALSE}
ticData <- pks$intensity/pks$normalizations$TIC
nonorm <- pks$intensity

#1000 de cada en bucle
nonormprueba2<-numeric()
for (h in 1:4)
{
  nonormprueba2 <-c(nonormprueba2,  pks$intensity[pks$region==new_names[h]][sample(length(pks$intensity[pks$region==new_names[h]]),1000)])
  
}

data_norm<-nonormprueba2

#normalized_data <- ticData

pca <- prcomp(data_norm, center = T, scale. = T)

plotpca <- rMSIproc::plotValuesImage(pks, pca$x[,1], labels = new_names, scale_title = "TIC PC1")
plotpca <- rMSIproc::plotValuesImage(pks, pca$x[,2], labels = new_names, scale_title = "TIC PC2")
plotpca <- rMSIproc::plotValuesImage(pks, pca$x[,3], labels = new_names, scale_title = "TIC PC3")

x_PC <- 1
y_PC <- 2
variX <- 100*((pca$sdev[x_PC])^2/sum(pca$sdev^2))
variY <- 100*((pca$sdev[y_PC])^2/sum(pca$sdev^2))
pltDF <- data.frame( xPC = pca$x[,x_PC], yPC = pca$x[,y_PC], sample = pks$region)
pltDFrandomized <- pltDF[sample(nrow(pltDF)), ] #Randomize rows to get a better plot

bi_1 <- ggplot( data = pltDFrandomized, aes( x = xPC, y = yPC))
bi_1_ <- bi_1 + theme_light()
bi_1 <- bi_1 + geom_point( aes( colour = sample), alpha = 0.5, size = 2)
bi_1 <- bi_1 + scale_x_continuous(name = paste0("PC", x_PC, " (", round(variX, digits = 2), "%)"))
bi_1 <- bi_1 + scale_y_continuous(name = paste0("PC", y_PC, " (", round(variY, digits = 2), "%)"))
bi_1 <- bi_1 + ggtitle(paste0("TIC PC", x_PC, " vs. PC", y_PC, " TIC CENTER SCALE TRUE"))
bi_1 + theme_light()

clus <- kmeans(ticData, 4)
mclors <- rMSIproc::plotClusterImage(pks, clus$cluster)